Improved state-space models for inference about spatial and temporal variation in abundance from count data

### 2012-10-02 Dail_Mad_ext.tex
future

The classical approach to computing $[N^+_t]$ is based upon Empirical
Bayes methods. Note that neither Empiriis not actually a Bayesian
method, it is used to estim.
We take the parameter estimates and plug them back in.... For example
to obtain the posterior distribution for $[N_{i1}]$
\begin{equation}
  [\hat{N}_{i1}] = \frac{[N_{i1}|\hat{\Lambda}][X_{i1}|N_{i1}, \hat{p}]}{\sum_{N_{i1}=0}^K [N_{i1}|\hat{\Lambda}][X_{i1}|N_{i1},\hat{p}]}
  \label{eq:3}
\end{equation}
Once we have estimated the posterior distribution for the first time
period, we can then proceed to estimate the posterior distribution for
subsequent time periods using recursion.

Note that when $X_{it}$ is not observed, including for future time
periods, the distribution $[X_{it}|N_{it}]$ is taken to be uniform
such that it has no influence on the posterior. In other words, the
posterior for years with no observed data is determined entirely by
the posterior for the previous year (and the population dynamics
parameters).

If covariates of the dynamics parameters are present, the problem
becomes much harder because, to fully account for uncertainty, we need
a model for the future values of the covariates as well.

In this case, the Bayesian approach is the most direct and easy to
implement because population size at any future point in time is
regarded as an unknown parameter like any other. Thus, one can simply
compute the posterior distribution of $N^+_t$, which is called the
posterior predictive distribution. In practical terms, this simply
means adding ``missing data'' to the observed count data, and allowing
the MCMC algorithm to sample the latent abundance variables
conditional only on abundance in each previous time period (and the
parameters governing population dynamics). An example of this is
provided in the appendix.

>>>>

The classical approach to computing $[N^+_t]$ is based upon Empirical
Bayes methods. Note that neither Empiriis not actually a Bayesian
method, it is used to estim.
We take the parameter estimates and plug them back in.... For example
to obtain the posterior distribution for $[N_{i1}]$
\begin{equation}
  [\hat{N}_{i1}] = \frac{[N_{i1}|\hat{\Lambda}][X_{i1}|N_{i1}, \hat{p}]}{\sum_{N_{i1}=0}^K [N_{i1}|\hat{\Lambda}][X_{i1}|N_{i1},\hat{p}]}
  \label{eq:eb1}
\end{equation}
Once we have estimated the posterior distribution for the first time
period, we can then proceed to estimate the posterior distribution for
subsequent time periods using recursion.
\begin{equation}
  [\hat{N}_{it}] =
  \frac{[N_{it}|N_{it-1},\hat{\bm{\Theta}}][X_{it}|N_{it},
    \hat{p}]}{\sum_{N_{it}=0}^K [N_{it}|N_{it-1}, \hat{\Theta}][X_{it}|N_{it},\hat{p}]}
  \label{eq:eb2}
\end{equation}

Note that when $X_{it}$ is not observed, including for future time
periods, the distribution $[X_{it}|N_{it}]$ is taken to be uniform
such that it has no influence on the posterior. In other words, the
posterior for years with no observed data is determined entirely by
the estimated posterior from the previous year and the population dynamics
parameters.

If covariates of the dynamics parameters are present, the problem
becomes much harder because, to fully account for uncertainty, we need
a model for the future values of the covariates as well. When the
covariates are continuous, this requires integrating over their
possible values, which becomes very challenging
computationally. Indeed, simulation methods such as MCMC are the only
viable options when numerous covariates are present, which will often
be the case in many ecological studies.

Relative to the computational challenges associated with the empirical
Bayes methods, the Bayesian approach is much more easy to implement,
especially since MCMC will be generally be relied upon in the first
place. Futhermore, because population size at any future point in time is
regarded as an unknown parameter like any other, computing the
posterior distribution of $N^+_t$ requires no additional steps in MCMC
algorithm, other than formally stating that the count data are ``missing''
in future years.  An example of this is provided in the appendix.
